---
title: "STA 445 Assignment 5"
author: "Kaylin McLiverty"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
library(patchwork)
library(viridis)
library(latex2exp)
library(plotly)
library(tidyverse)
library(ggrepel)
```

## Exercise 1
The `infmort` data set from the package `faraway` gives the infant mortality
    rate for a variety of countries. The information is relatively out of date
    (from 1970s?), but will be fun to graph. Visualize the data using by creating
    scatter plots of mortality vs income while faceting using `region` and setting
    color by `oil` export status. Utilize a $\log_{10}$ transformation for both
    `mortality` and `income` axes. This can be done either by doing the
    transformation inside the `aes()` command or by utilizing the `scale_x_log10()`
    or `scale_y_log10()` layers. The critical difference is if the scales are on
    the original vs log transformed scale. Experiment with both and see which you
    prefer.

    a)  The `rownames()` of the table gives the country names and you should
        create a new column that contains the country names. *`rownames`
```{r}
data(infmort, package = "faraway")

infmort1 <- infmort %>% mutate( country = rownames(infmort) )
head(infmort1)
```
        
    b)  Create scatter plots with the `log10()` transformation inside the `aes()`
        command.
        
```{r}
P1 <- ggplot(infmort1, aes(x=log10(mortality), y=log10(income))) + 
  geom_point(aes(color=oil)) +
  facet_wrap( ~region )+
  scale_color_brewer(palette='Dark2')
P1
```

    c)  Create the scatter plots using the `scale_x_log10()` and `scale_y_log10()`.
        Set the major and minor breaks to be useful and aesthetically pleasing.
        Comment on which version you find easier to read.
        
*I personally think that the version without the 'scale_x_log10()' is easier to read.*
```{r}
P2 <- ggplot(infmort1, aes(x=log10(mortality), y=log10(income))) + 
  geom_point(aes(color=oil)) +
  facet_wrap( ~region )+
  scale_color_brewer(palette='Dark2')+
   scale_x_log10(breaks=(0:3), 
                minor_breaks=seq(0,3, by=.5))+
  xlab('Mortality (1000s of infants)')+
  scale_y_log10(breaks=(0:4), 
                minor_breaks=seq(0,4, by=.35))+
  ylab('Income (1000s of dollars)')
P2
```
        
    d)  The package `ggrepel` contains functions `geom_text_repel()` and
        `geom_label_repel()` that mimic the basic `geom_text()` and `geom_label()`
        functions in `ggplot2`, but work to make sure the labels don't overlap.
        Select 10-15 countries to label and do so using the `geom_text_repel()`
        function.

```{r}
P3 <- ggplot(infmort1, aes(x=log10(mortality), y=log10(income))) + 
  geom_point(aes(color=oil)) +
  facet_wrap( ~region )+
  scale_color_brewer(palette='Dark2')+
  scale_x_log10(breaks=(0:3), 
                minor_breaks=seq(0,3, by=.5))+
  xlab('Mortality (1000s of infants)')+
  scale_y_log10(breaks=(0:4), 
                minor_breaks=seq(0,4, by=.35))+
  ylab('Income (1000s of dollars)')+
  geom_text_repel(aes(label=country))
P3
```
        

## Exercise 2
Using the `datasets::trees` data, complete the following:

    a)  Create a regression model for $y=$ `Volume` as a function of $x=$ `Height`.
```{r}
data(trees)

model <- lm( Volume ~ Height, data=trees)
```
    
    b)  Using the `summary` command, get the y-intercept and slope of the
        regression line.
```{r}
summary(model)

trees <- trees %>% mutate(fit=fitted(model))
```
        
    c)  Using `ggplot2`, create a scatter plot of Volume vs Height. 
```{r}
P4 <- ggplot(trees, aes(x=Height, y=Volume)) +
  geom_point()
P4
```
    
    d)  Create a nice white filled rectangle to add text information to using by
        adding the following annotation layer.
```{r, eval=FALSE}
P4 <- ggplot(trees, aes(x=Height, y=Volume)) +
        annotate('rect', xmin=65, xmax=75, ymin=60, ymax=74,
                 fill='white', color='black') +
        geom_point()
P4
```
    e)  Add some annotation text to write the equation of the line
        $\hat{y}_i = -87.12 + 1.54 * x_i$ in the text area.
```{r}
P4 <- ggplot(trees, aes(x=Height, y=Volume)) +
        annotate('rect', xmin=65, xmax=75, ymin=60, ymax=74,
                 fill='white', color='black') +
        annotate('text', x=70, y=68, 
           label=latex2exp::TeX('$\\hat{y}_i = -87.12 + 1.54 * x_i$') ) +
        geom_point()
P4
```
        
    f)  Add annotation to add $R^2 = 0.358$
```{r}
P4 <- ggplot(trees, aes(x=Height, y=Volume)) +
        annotate('rect', xmin=65, xmax=75, ymin=60, ymax=74,
                 fill='white', color='black') +
        annotate('text', x=70, y=70, 
           label=latex2exp::TeX('$\\hat{y}_i = -87.12 + 1.54 * x_i$') ) +
        annotate('text', x=70, y=65,label=latex2exp::TeX('$R^2 = 0.358$'))+
        geom_point()
P4
```
    
    g)  Add the regression line in red. The most convenient layer function to uses
        is `geom_abline()`. It appears that the `annotate` doesn't work with
        `geom_abline()` so you'll have to call it directly.
```{r}
P4 <- ggplot(trees, aes(x=Height, y=Volume)) +
        annotate('rect', xmin=65, xmax=75, ymin=60, ymax=74,
                 fill='white', color='black') +
        annotate('text', x=70, y=70, 
           label=latex2exp::TeX('$\\hat{y}_i = -87.12 + 1.54 * x_i$') ) +
        annotate('text', x=70, y=65,label=latex2exp::TeX('$R^2 = 0.358$'))+
        geom_point()+
        geom_abline(intercept=-87.1236, slope=1.5433, color='red')

P4
```



  *I wanted to compare the two graphs. The first one uses geom_abline and the second uses geom_line with the column of fitted values that I mutated* 
  
  
```{r}
P4 <- ggplot(trees, aes(x=Height, y=Volume)) +
        annotate('rect', xmin=65, xmax=75, ymin=60, ymax=74,
                 fill='white', color='black') +
        annotate('text', x=70, y=70, 
           label=latex2exp::TeX('$\\hat{y}_i = -87.12 + 1.54 * x_i$') ) +
        annotate('text', x=70, y=65,label=latex2exp::TeX('$R^2 = 0.358$'))+
        geom_point()+
        geom_line( aes(y=fit), color='red')

P4
```
        
        